---
title: "Functions required for sampling of nonpresence data"
author: "EZD Team"
date: "20 May 2019"
output: 
  html_document:
    toc: true
    toc_depth: 3
    toc_float: true
    number_sections: true
    df_print: paged
    code_folding: hide
---

```{r setup, include=FALSE, echo = FALSE}
knitr::opts_chunk$set(collapse = TRUE, 
                      warning = FALSE, 
                      message = FALSE, 
                      fig.width = 14, dpi = 300)

# Load Packages
source("loadpackages.R")
```

# Loading all required objects

- Shapefiles for rivers and Loess
```{r eval=FALSE}
shapewater <- shapefile("../../Daten/Raster/RawFiles/DEU_water_lines_dcw.shx")
loess_shape <- shapefile("../../Daten/Raster/RawFiles/Loess/gk500_haupteinheiten_epsg4258.shx")
```

- Rasterfiles that are required for sampling points from within Bavaria
```{r eval=FALSE}
ger <- raster::getData(name = "GADM", country = "Germany", level = 1)
ger_alt <- raster::getData(name = "alt", country = "Germany")
```

- The coordinates of Bavarias border will serve as a default border for bavSamp()
```{r eval=FALSE}
bayern_coords <- ger@polygons[[2]]@Polygons[[1]]@coords
```

# Automating Samplingprocess
- Function to sample random points from bavaria

```{r eval=FALSE}
#' bavSamp: Sample random points from within the borders of Bavaria. The Function will run a set amount of iterations until returning a spatial.points.data.frame that contains a random sample of coordinate pairs within Bavaria. The function initializes an empty spatial.points.data.frame that will gradually be filled up with the sampled points. 
#'
#' @param from Raster from which the points are sampled originally.
#' @param within Polygon, borders of region in which the points should fall.
#' @param size Size of coordinate pair sample that gets drawn in every iteration. 
#' @param maxiter Maximum number of iterations the function will run.
#' @param rlength Number of points that should be returned.
#'
#' @return a spatial.points.data.frame
#' @export
#'
#' @examples
bavSamp <- function(from = ger_alt, within = bayern_coords, size = 1000, maxiter = 100, rlength = 100){
  i <- 0
  iteration <- 1
  returnframe <- new("SpatialPoints",   
                     coords = structure(numeric(0), .Dim = c(0L, 2L),                                  .Dimnames = list(NULL, c("coords.x1", "coords.x2"))),  
                     bbox = structure(c(1, 1, 1, 1), .Dim = c(2L, 2L),                                 .Dimnames = list(c("coords.x1", "coords.x2"), c("min", "max"))),
                     proj4string = new("CRS", projargs = NA_character_))
  if (i >= maxiter) {
    warning("Maximum number of iterations reached \n Returning sample.")
    warning("Sampling iterations: ", print(i))
    return(returnframe)    
  }
  while (i < maxiter) {
    print(iteration)
    temp <- sampleRandom(from, size, sp = T)
    checkvector <- point.in.polygon(temp@coords[,1], temp@coords[,2], bayern_coords[,1], bayern_coords[,2])
    checkvector <- as.logical(checkvector)
    temp <- temp[checkvector,]
    returnframe <- union(returnframe, temp)
    iteration <- iteration + 1
    if (nrow(returnframe@coords) >= rlength) {
      returnframe <- returnframe[1:rlength,]
      # set i equal to maxiter to abort sampling process early
      i <- maxiter
    }
  }
  return(returnframe)
}
```
- Function to split data.frames into parts of equal length.
```{r eval=FALSE}
#' Splitframe: Split data.frame into parts of equal length.
#'
#' @param df Input data.frame.
#' @param splits Number of parts to split the data.frame into. Nrow of input data.frame has to be a multiple of splits.
#'
#' @return A list of the split data.frames. Output in the form of a list is here desired, since this function is being called in order to utilize parallelization supplied by the foreach and %dopar% operators. 
#' @export
#'
#' @examples
splitframe <- function(df, splits) {
  if (mod(nrow(df), splits) != 0) {
    stop("nrow is not a multiple of splits")
  }
  pointsliste <- list()
  j <- 0
  for (i in 1:splits) {
    pointsliste[[i]] <- df[(j + 1):(i*(nrow(df)/splits)),]
    j <- j + (nrow(df)/splits)
  }
  return(pointsliste)
}
```
- Function to calculate accurate point distances to the rivershapefile. 
```{r eval=FALSE}
#' pardist2water: Calculating point distances to rivers in parallel. The approach via the foreach and doParallel libraries might be less elegant than the builtin functionality for parallel calculation supplied by R on linux systems, but this code should run independent of operating systems.
#'
#' @param pnts A data.frame containing coordinates.
#' @param shapef The rivershapefile the distances will be calculated to
#' @param splits The amount of splits the splitframe() call will execute.
#' @param workers The amount of workers for parallel computation. For 4 core systems 4 workers seem ideal to reduce overhead by setting up more workers for every thread. For larger datasets or stronger processors a larger workforce would of course be ideal. 
#'
#' @return A dataframe augmented with the desired distances.
#' @export
#'
#' @examples
pardist2water <- function(pnts, shapef = shapewater, splits = 10, workers = 4) {
  pnts <- as.data.frame(pnts)
  # split into equal parts
  pointsliste <- splitframe(pnts, splits)
  registerDoParallel(cores = workers)
  tmp <- foreach(i = 1:splits) %dopar% {
    geosphere::dist2Line(p = pointsliste[[i]], line = shapef)
  }
  # unlist tmp to data.frame
  tmp <- do.call(rbind, tmp)
  # bind results with original points
  tmp <- cbind(pnts, tmp)
  return(tmp)
}
```
- Function to fully automate sampling process for coordinate pairs
```{r eval=FALSE}
#' runValidation: Full automated point sampling and point distance calculation.
#'
#' @param numrows Number of desired points.
#' @param rseed Random seed to make results replicable.
#'
#' @return A data.frame.
#' @export
#'
#' @examples
runValidation <- function(numrows, rseed = 1){
  set.seed(rseed)
  tmp.set <- bavSamp(size = 5000, rlength = numrows)
  tmp.points <- tmp.set@coords
  tmp.points <- as.data.frame(tmp.points)
  tmp.points <- pardist2water(tmp.points, splits = 8, workers = 8)
  write.table(tmp.points, file = "../../Daten/negativdaten50k.csv")
  return(tmp.points)
}
```
# Filtering nonsites
- Importing data sets to work with
```{r eval=FALSE}
sitesdf <- read.table(file = "../../Daten/fenderdistance2water.csv")
nonsitesdf <- read.table(file = "../../Daten/negativdaten50k.csv")
```

- Distance function that will be used to calculate point distances
```{r eval=FALSE}
#' Haversine distance corrects for the curvature of the earth and thus most accurately calculates the distances between two given points 1 and 2. It also runs faster than the euclidian counterpart. 
#'
#' @param lon1 Longitude of point 1.
#' @param lat1 Latitude of point 1.
#' @param lon2 Longitude of point 2.
#' @param lat2 Longitutde of point 2.
#' @param r The radius of the sphere the points are located on. Default case equals the radius of the earth.
#'
#' @return The calculated distance.
#' @export
#'
#' @examples
dist_haversine <- function(lon1, lat1, lon2, lat2, r = 6378.388){
  # r = radius of the sphere the coordinates are located on e.g. the earth
  # calculate distance and return result in km
  temp <- 2*r*asin(sqrt((sin(((lat2 - lat1)*(pi/180))/2))^2 + cos(lat1*(pi/180))*cos(lat2*(pi/180)) * sin((((lon1 - lon2)*(pi/180))/2))^2))
  return(temp)
}
```
- Filter function to remove all points that are too close to the sites.
```{r eval=FALSE}
#' Filters nonsite points that are too close to sites. 
#'
#' @param samplefrom Data.frame that contains nonsites in question.
#' @param checkagainst Data.frame that contains sites to be checked.
#' @param margin Radius in km around sites to remove nonsites from. 
#'
#' @return The samplefrom data.frame without the filtered points.
#' @export
#'
#' @examples
filterbymargin_haversine <- function(samplefrom, checkagainst, margin = 1.5) {
  df <- samplefrom
  df$temporary <- 0
  for (i in 1:nrow(checkagainst)){
    df$temporary <- mapply(dist_haversine, df$lon, df$lat, checkagainst$lon[i], checkagainst$lat[i])
    df <- df[(df$temporary>margin),]
  }
  df$temporary <- NULL
  return(df)
}
```
- Filtering out nonsites that are close than 1.5 km to sites
```{r eval=FALSE}
negative50kfiltered <- filterbymargin_haversine(samplefrom = nonsitesdf, checkagainst = sitesdf, margin = 1.5)
nrow(negative50kfiltered)
write.table(negative50kfiltered, file = "../../Daten/neg29kfiltered.csv")
```



