---
title: "Generation of Environmental Raster Data"
author: "EZD Team"
date: "20 May 2019"
output: 
  html_document:
    toc: true
    toc_depth: 3
    toc_float: true
    number_sections: true
    df_print: paged
    code_folding: hide
---

```{r}
'Evtl adden:
density(eval_maxent_v) etc


'

```


```{r setup, include=FALSE, echo = FALSE}
knitr::opts_chunk$set(collapse = TRUE, 
                      warning = FALSE, 
                      message = FALSE, 
                      fig.width = 14, dpi = 300)

# Load Packages
library(dismo)
library(raster)
library(dplyr)
```


# DELETE
```{r}
predictors <- stack(c(
  "../../Daten/Raster/height_raster.grd",
  "../../Daten/Raster/temp_raster.grd",
  "../../Daten/Raster/rain_raster.grd",
  "../../Daten/Raster/distance_water_raster.grd",
  "../../Daten/Raster/distance_loess_raster.grd",
  "../../Daten/Raster/frostdays_raster.grd",
  "../../Daten/Raster/sunhours_raster.grd",
  "../../Daten/Raster/corine_raster.grd",
  "../../Daten/Raster/tpi_raster.grd",
  "../../Daten/Raster/slope_raster.grd",
  "../../Daten/Raster/aspect_raster.grd"))

predictors <- stack(c(
  predictors[[1]]/100,
  predictors[[2]],
  predictors[[3]],
  predictors[[4]]/1000,
  predictors[[5]]/1000,
  predictors[[6]],
  predictors[[7]],
  predictors[[8]],
  predictors[[9]],
  predictors[[10]],
  predictors[[11]]
  ))
```





# MaxEnt

## Presence Data

- Load Data and Change Coordinates to Numeric
```{r}
presence <- readRDS("../../Daten/Daten_mit_Epoche.RDS")
presence$lng_wgs84 <- as.numeric(as.character(presence$lng_wgs84))
presence$lat_wgs84 <- as.numeric(as.character(presence$lat_wgs84))
```

- Filter for Eisenzeit and Type and Select Coordinates
```{r}
# All Types
presence_all <- filter(presence, Epoche %in% c("Latènezeit", "Hallstattzeit"))
presence_all <- presence_all[6:7]

# Siedlung
presence_s <- filter(presence, Siedlung == TRUE)
presence_s <- presence_s[6:7]

# Viereckschanze
presence_v <- filter(presence, Typ == "Viereckschanze")
presence_v <- presence_v[6:7]
```

## Preparation for MaxEnt

- Split into Training and Testing Set
```{r}
set.seed(12345)

# All Types
group <- kfold(presence_all, 5)
presence_train_all <- presence_all[group != 1, ]
presence_test_all <- presence_all[group == 1, ]

# Siedlung
group <- kfold(presence_s, 5)
presence_train_s <- presence_s[group != 1, ]
presence_test_s <- presence_s[group == 1, ]

# Viereckschanze
group <- kfold(presence_v, 5)
presence_train_v <- presence_v[group != 1, ]
presence_test_v <- presence_v[group == 1, ]
```

## MaxEnt Model

### All Types

- Fit Model
```{r}
model_maxent_all <- maxent(predictors, presence_train_all,
                           factors = c("corine", "aspect"), 
                           nbg = nrow(presence_train_all), remove.duplicates = T, 
                           args = c("responsecurves"))#,
                           #path = "Maxent")
```

- Variable Importance
```{r}
plot(model_maxent_all)
```

- Density
```{r}
#density(model_maxent_all)
```

- Response Curve
```{r}
response(model_maxent_all)
```

- Prediction
```{r}
pred_maxent_all <- predict(model_maxent_all, predictors)
plot(pred_maxent_all)
print(pred_maxent_all)
```

- Evaluation
```{r}
bg <- randomPoints(predictors, nrow(presence_test_all))
eval_maxent_all <- evaluate(model_maxent_all, p = presence_test_all, a = bg, x = predictors)
threshold(eval_maxent_all)
plot(eval_maxent_all, "ROC")
boxplot(eval_maxent_all)
print(eval_maxent_all)
```

### Siedlung

- Fit Model
```{r}
model_maxent_s <- maxent(predictors, presence_train_s,
                           factors = c("corine", "aspect"), 
                           nbg = nrow(presence_train_s), remove.duplicates = T, 
                           args = c("responsecurves"))#,
                           #path = "Maxent")
```

- Variable Importance
```{r}
plot(model_maxent_s)
```

- Density
```{r}
#density(model_maxent_s)
```

- Response Curve
```{r}
response(model_maxent_s)
```

- Prediction
```{r}
pred_maxent_s <- predict(model_maxent_s, predictors)
plot(pred_maxent_s)
print(pred_maxent_s)
```

- Evaluation
```{r}
bg <- randomPoints(predictors, nrow(presence_test_s))
eval_maxent_s <- evaluate(model_maxent_s, p = presence_test_s, a = bg, x = predictors)
threshold(eval_maxent_s)
plot(eval_maxent_s, "ROC")
boxplot(eval_maxent_s)
print(eval_maxent_s)
```

### Viereckschanze

- Fit Model
```{r}
model_maxent_v <- maxent(predictors, presence_train_v,
                           factors = c("corine", "aspect"), 
                           nbg = nrow(presence_train_v), remove.duplicates = T, 
                           args = c("responsecurves"))#,
                           #path = "Maxent")
```

- Variable Importance
```{r}
plot(model_maxent_v)
```

- Density
```{r}
#density(model_maxent_v)
```

- Response Curve
```{r}
response(model_maxent_v)
```

- Prediction
```{r}
pred_maxent_v <- predict(model_maxent_v, predictors)
plot(pred_maxent_v)
print(pred_maxent_v)
```

- Evaluation
```{r}
bg <- randomPoints(predictors, nrow(presence_test_v))
eval_maxent_v <- evaluate(model_maxent_v, p = presence_test_v, a = bg, x = predictors)
threshold(eval_maxent_v)
plot(eval_maxent_v, "ROC")
boxplot(eval_maxent_v)
print(eval_maxent_v)
```







